\Title

### 2021-06-01 latex/paper/main.tex
Continued writing the proof of Theorem.

\begin{lemma}\label{lemma:energy-regulation-D}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    parameterized by \((q_u,p_u) \in \SxR\).
    Let \(D \subset \SxR\) be open and 
    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    satisfying \(\dot{r} \equiv 0\) and \(\dot{\theta} > 0\) when
    \(I = 0\).
    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
    \((\partial g/\partial I)(\theta,r,0) \equiv 0\).
    Define \(b(\theta,r_0) = (\partial g/\partial r)(\theta,r_0,0)\);
    assume it is \(2\pi\)-periodic in \(\theta\) for fixed \(r_0\),
    and that there exists \(\epsilon > 0\) such that 
    \[
        \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
        ,
    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
    for all \(I \in [-I^\star, 0[\).
\end{lemma}
\begin{proof}
    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, ]0,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE 
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \frac{\dot{r}}{\dot{\theta}} =
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    where the prime denotes differentiation by \(\theta\) and 
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
    Taking the Taylor expansion of this solution around \(I = 0\) we get
    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
    + R(\theta,r_0,I)\). 
    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) for all \(\theta\),
    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
    \begin{equation}
        \begin{cases}
        \hat{r}_1^\prime = 
        \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
        \frac{\partial g}{\partial I}(\theta,r_0,0)
        , \\
        \hat{r}_1(0) = 0
        ,
    \end{cases}
    \end{equation}
    and the remainder \(R(\theta,r_0,I)\) is \(O(I^2)\).

    By assumption, \((\partial g/\partial I)(\theta,r_0,0) \equiv 0\) and
    \(b(\theta,r_0) := (\partial g/\partial \hat{r})(\theta,r_0,0)\). 
    Hence, one can compute
    \[
        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
        .
    \]
    Since and \(\dot{\theta} > 0\) and \(b(\theta,r_0)\) is \(2\pi\)-periodic in
    \(\theta\), there is a well-defined Poincar\'{e} map describing
    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\);
    this map is 
    \begin{align*}
        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
        .
    \end{align*}
    This map expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\), which
    means we can understand the evolution of \(\hat{r}\) by studying
    the dynamics of the discrete-time system
    \begin{equation}\label{eqn:r-discrete}
        r_{n+1} = r_n + IS(r_n) + R(2\pi,r_n,I)
        ,
    \end{equation}
    with initial condition \(r_0\).
    Here, \(r_n\) corresponds to the distance along the line segment
    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
    when an orbit of the constrained dynamics hits \(L\) for the
    the \(n^\text{th}\) time, assuming the orbit was initialized at
    \((r,\theta) = (r_0,0)\).

    Proving the constrained dynamics gain energy on \(D\) is equivalent to
    proving \(\hat{r}_{n}\) eventually reaches \(\overline{r}\).
    Thanks to a perturbation analysis from \cite[Theorem 10.1]{khalil_nonlinear}, 
    we know there exists some \(I_2 \in ]0,I_1]\) and \(\delta > 0\)
    such that \(\left|R(2\pi,r_n,I)\right| \leq \delta I^2\) for all 
    \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0)\) is uniformly
    bounded below by \(\epsilon\), we can find a lower bound on
    \eqref{eqn:r-discrete}:
    \[
        r_{n+1} \geq r_n + I\epsilon - \delta I_2^2
        .
    \]
    Choosing \(I^\star \in ]0,I_2]\) small enough makes 
    \(I\epsilon - \delta I_2^2 \geq \gamma > 0\) for all \(I \in ]0,I^\star]\).
    Therefore, \(r_{n+1} \geq r_n + \gamma \geq r_0 + n\gamma\), which means all
    solutions of \eqref{eqn:r-discrete} flow towards \(\overline{r}\).
    We conclude that, for \(I \in ]0, I^\star]\), the constrained dynamics gain
    energy on \(D\).
    Changing the sign of \(I\) means \(r_{n+1} \leq r_n - n\gamma\), in which
    case the constrained dynamics lose energy on \(D\).
\end{proof}

\textbf{TODO: Prove origin is unstable, then apply lemma to O1 and O2 and show
that orbits go through Epi}
%/========== Conclusion ==========/%
\section{Conclusion}\label{sec:conclusion}

>>>>

\begin{lemma}\label{lemma:energy-regulation-D}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    parameterized by \((q_u,p_u) \in \SxR\).
    Let \(D \subset \SxR\) be open and 
    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    satisfying \(\dot{r} \equiv 0\) and \(\dot{\theta} > 0\) when
    \(I = 0\).
    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
    \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
    Define \(b(\theta,r_0) = (\partial g/\partial I)(\theta,r_0,0)\);
    assume it is \(2\pi\)-periodic in \(\theta\) for any fixed \(r_0\),
    and that there exists \(\epsilon > 0\) such that 
    \[
        S(r_0) := \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
        ,
    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
    for all \(I \in [-I^\star, 0[\).
\end{lemma}
\begin{proof}
    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, ]0,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE 
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \frac{\dot{r}}{\dot{\theta}} =
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    where the prime denotes differentiation by \(\theta\) and 
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
    Taking the Taylor expansion of this solution around \(I = 0\) we get
    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
    + R(\theta,r_0,I)\). 
    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) for all \(\theta\),
    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
    \begin{equation}\label{eqn:r1-hat-prime}
        \begin{cases}
        \hat{r}_1^\prime = 
        \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
        \frac{\partial g}{\partial I}(\theta,r_0,0)
        , \\
        \hat{r}_1(0) = 0
        ,
    \end{cases}
    \end{equation}
    and the remainder \(R(\theta,r_0,I)\) is \(O(I^2)\).

    Our assumption that \((\partial g/\partial r)(\theta,r_0,0) \equiv 0\)
    implies that \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
    The solution of \ref{eqn:r1-hat-prime} can be obtained by quadrature, giving
    \[
        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
        ,
    \]
    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
    Since and \(\dot{\theta} > 0\) and \(b(\theta,r_0)\) is \(2\pi\)-periodic in
    \(\theta\), there is a well-defined Poincar\'{e} map describing
    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
    This map is 
    \begin{align*}
        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
        ,
    \end{align*}
    which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
    Hence, we can understand the evolution of \(\hat{r}\) by studying
    the dynamics of the discrete-time system
    \begin{equation}\label{eqn:r-discrete}
        r_{n+1} = r_n + IS(r_n) + R(2\pi,r_n,I)
        ,
    \end{equation}
    with initial condition \(r_0\).
    Here, \(r_n\) corresponds to the distance along the line segment
    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
    assuming the orbit was initialized at \((r,\theta) = (r_0,0)\).

    Proving the constrained dynamics gain energy on \(D\) is equivalent to
    proving \(\hat{r}_{n}\) eventually reaches \(\overline{r}\).
    Thanks to \cite[Theorem 10.1]{khalil_nonlinear}, 
    we know there exists some \(I_2 \in ]0,I_1]\) and \(\delta > 0\)
    such that \(\left|R(2\pi,r_n,I)\right| \leq \delta I^2\) for all 
    \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0)\) is uniformly
    bounded below by \(\epsilon\), we can find a lower bound on
    \eqref{eqn:r-discrete}:
    \[
        r_{n+1} \geq r_n + I\epsilon - \delta I_2^2
        .
    \]
    Choosing \(I^\star \in ]0,I_2]\) small enough makes 
    \(I\epsilon - \delta I_2^2 \geq \gamma > 0\) for all \(I \in ]0,I^\star]\).
    Therefore, \(r_{n+1} \geq r_n + \gamma \geq r_0 + n\gamma\), which means all
    solutions of \eqref{eqn:r-discrete} flow towards \(\overline{r}\).
    We conclude that, for \(I \in ]0, I^\star]\), the constrained dynamics gain
    energy on \(D\).
    Changing the sign of \(I\) means \(r_{n+1} \leq r_n - n\gamma\), in which
    case the constrained dynamics lose energy on \(D\).
\end{proof}

We will now prove the first part of Theorem
\ref{thm:acrobot-energy-stabilization}.
To apply Lemma \ref{lemma:energy-regulation-D}, we choose \(D\) to be the
annulus \(\mathcal{O}_1 \backslash \{(0,0)\}\) consisting of all oscillations of
the nominal pendulum.
Using the procedure described by Mohammadi \etal 
\cite{dynamic_vhcs_stabilize_closed_orbits} for finding pseudo-polar
coordinates adapted to level sets of \(E\), we compute
\begin{align}
    \label{eqn:r-osc}
    r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right)
    , \\
    \label{eqn:theta-osc}
    \theta &=
    \left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
    \right|_{r = r(q_u,p_u)}
    ,
\end{align}
where \(r \in \, ]0,\pi[\).
We now verify these coordinates satisfy the requirements of 
Lemma \ref{lemma:energy-regulation-D}.
When \(I = 0\) we have \(\dot{r} = 0\) and
\begin{equation} \label{eqn:theta-dot-nom-osc}
    \dot{\theta} = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(r\cos(\theta)) - \cos(r)}
            {r^2 \sin(\theta)^2}}
    .
\end{equation}
Note that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
\(\theta \in \{0,\pi\}\).
Taking the limit as \(\theta\) approaches these points yields
\begin{equation*}
    \lim\limits_{\theta \to 0}\dot{\theta} 
    = \lim\limits_{\theta \to \pi} \dot{\theta}
    = \sqrt{\frac{6g \sin(r)}{10lr}}
    , 
\end{equation*}
which is smooth and well-defined for all
\(r \in \, ]0,\pi[\).
Taking these removable singularities into account,
one can verify that \(\dot{\theta} > 0\) for every \(r \in \, ]0,\pi[\).
Setting \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.
Next, we compute
\begin{equation*}
    b(\theta,r_0) = 
    \frac{\partial g}{\partial I}(\theta, r_0, 0) = K a(\theta,r_0)
    ,
\end{equation*}
where
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(\theta,r_0) &:= \frac{
        r_0 |\sin(\theta)| \left(
        5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
    \right)
    }{
    \sin(r_0)\sqrt{\cos(r_0c_\theta) - c_{r_0}}
    }
    .
\end{align*}
Notice that \(a(\theta,r_0)\) does not depend on the physical parameters of the
acrobot.

\textbf{TODO: write about Qro, show the figure, apply lemma to D,
    prove origin is unstable, show orbits will go through Epi, 
    then derive the constrained dynamics for the rotation domain and apply lemma
    there, concluding that O2 is O1 U Epi U rotation domain and is therefore
    also gaining energy.}

%/========== Conclusion ==========/%
\section{Conclusion}\label{sec:conclusion}
